Navigating Hostility: The Effect of Nonprofit Transparency and Accountability on Donor Preferences in the Face of Shrinking Civic Space

### 2025-04-15 manuscript/manuscript.qmd
Remove this old profile stuff
Final edits pre-final submission

---
title: "Navigating Hostility: The Effect of Nonprofit Transparency and Accountability on Donor Preferences in the Face of Shrinking Civic Space"
short-title: "Navigating Hostility"
date: "2025-01-22"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
correspondence-prefix: "Correspondence concerning this article should be addressed to"
abstract: |
  Governments across the world have increasingly used laws to restrict the work of nonprofits, which has led to a reduction in public or official foreign aid directed towards these groups. Many international nonprofits, in response, have turned to individual donors to offset the loss of traditional funding. What are individual donors' preferences regarding donating to legally besieged nonprofits abroad? We conducted a conjoint experiment on a nationally representative sample of likely donors in the US and found that learning about host government criticism and legal restrictions on nonprofits decreases individuals' preferences to donate to them. However, organizational features such as financial transparency and accountability can protect against this dampening effect. Our results have important implications both for understanding private international philanthropy and how nonprofits can better frame their fundraising appeals at a time when they are facing restrictive civic spaces and hostile governments abroad.
additional-info: |
  **Data Availability:** All data and replication code are publicly available at <https://github.com/andrewheiss/silent-skywalk>
keywords:
  - philanthropy
  - conjoint experiments
  - donor heuristics
  - repression
  - NGOs
  - civil society
  - nonprofits
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
word-count: "10,149"  # Subtract 77 for the final statements at the end
---

@@

tar_load(c(grid_treatment_only, level_lookup, feature_lookup))

@@

We fielded a conjoint experiment to a nationally representative sample of 1,016 adults in the US. This experiment allows us to assess individual-level giving preferences among likely donors in the US towards NGOs abroad working in a variety of issue areas---human rights, environment, emergency response, and refugee relief.[^not-elite] More importantly, it allows us to directly compare the relative impact of host government restriction of NGOs and organizational financial transparency and accountability on individual donor preferences. The result of this experiment shows that both financial transparency and accountability increase individuals' willingness to donate to NGOs. Conversely, learning about host government criticism of and restrictions on NGOs decreases the likelihood of donation by itself; however, organizational features such as financial transparency and accountability can protect against this dampening effect.

@@

A conjoint research design also allows us to estimate treatment effects even if respondents do not see every combination of organization attributes, and correspondingly provides a substantial increase in statistical power. In standard factorial experimental designs, participants would ordinarily need to be shown all possible combinations of experimental attributes. In our experiment, we presented respondents with six possible treatments with randomized attributes for each treatment, yielding 576 possible unique combinations of features. Even with respondents answering twelve iterations of the experiment, not every combination was seen. However, as long as all the possible organizational attributes are well randomized and there are no systematic biases toward specific options (i.e., more respondents select the first option because it is the first) or toward earlier iterations of attribute choices (i.e., respondents are more careful and attentive for the first hypothetical organization than the last), we can pool all observations together for specific attributes of interest while marginalizing across all other attributes [@KertzerRenshonYarhi-Milo:2021]. This allows us to (1) estimate the effect of each experimental treatment even if some unique combinations of attributes were unseen, and (2) use a much smaller sample size than would be required in a more traditional factorial design.

While the ability to select key attributes and marginalize over others provides us with analytic flexibility, it also raises possible issues with multiple comparisons, *p*-hacking, and selective cherry-picking [@BansakHainmuellerHopkins:2021], especially given the fact that we have 576 possible combinations of independent variables to explore. As such, before launching the survey experiment, we preregistered a subset of confirmatory and exploratory hypotheses at the Open Science Framework. These hypotheses deal specifically with our key research questions about the effect of transparency, accountability, and crackdown on the propensity to donate, along with comparison treatment effects for organization brand name, issue area, and funding sources. We also examine the interaction between transparency, accountability, and host-country relationships. Our preregistration protocol is available at `ANONYMIZED_URL` and in the appendix, and our data and reproducible code is available at `ANONYMIZED_URL`.[^prereg]

[^prereg]: For clarity, we focus on our simplest, non-nested hypotheses, omitting many preregistered ones that overlap (e.g., the effect of branding and issue area and transparency and crackdown simultaneously). Following @WillrothAtherton:2024, we outline all the deviations from our preregistration protocol in the appendix.

@@

We fielded our survey experiment through Centiment, which recruits representative samples of paid (and highly engaged) survey participants online. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question about their philanthropic behavior early in the survey---if a participant responded that they give once every few years or never, they were disqualified from the study. We also included an attention check question early in the survey and removed respondents who failed the question. Importantly, these screening questions were presented *prior to* the experimental manipulation to avoid post-treatment bias [@MontgomeryNyhanTorres:2018a]. Following screening, we received 1,016 viable responses. Tables A1, A2, and A3 provide a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across all pre-treatment characteristics, including gender, age, education, income, and attitudes toward charity. Additionally, Table A4 compares sample characteristics with nationally representative estimates from the 2019 Current Population Survey (CPS). Our sample matches national proportions of age, marital status, and education, and contains slightly more men and slightly higher income levels than the general population. Due to the initial screening question that limited participants to those with an interest in charitable activity, respondents are far more likely to have donated, volunteered, or voted in the past year than population CPS estimates. Our reported causal effects are therefore not representative of the entire population and instead are generalizable to people interested in charitable giving.

@@

At the time of designing the study, Charity Navigator---a large database of US nonprofit financial information---categorized international nonprofit activities into four general causes: (1) development, (2) humanitarian relief, (3) international affairs and human rights advocacy, and (4) environment.[^cn] Accordingly, we used four nonprofits that are stereotypical for each cause: the International Committee of the Red Cross (humanitarian relief), Amnesty International (human rights), Oxfam (development), and Greenpeace (environment).[^real-names] We also varied several other features, including four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), two organizational practices (financial transparency and third-party accountability audits), three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and three relationships with host governments (friendly, criticized by the government, and under government crackdown).

[^cn]: See <http://web.archive.org/web/20170315000746/https://www.charitynavigator.org/index.cfm?bay=search.categories&categoryid=7>. Under the explicitly "International" category, Charity Navigator listed Development and Relief Services; International Peace, Security, and Affairs; and Humanitarian Relief Supplies. The "Environment" category is listed separately but contains many INGOs such as Greenpeace and the World Wide Fund for Nature (WWF). At the time of writing, Charity Navigator had changed its taxonomy to include 18 broader categories.

@@

After presenting respondents with a set of three randomly generated organizations, we asked them which of the three they would be willing to donate to, along with an option for no selection (see @tbl-example for an example). This choice indicated donation intention and reflected the option the respondent preferred most within that set. We then repeated the process eleven more times for a total of twelve randomized iterations of hypothetical combinations of attributes, resulting in 12,192 completed experimental tasks (12 iterations × 1,016 respondents). This iterative process across twelve experimental tasks is typical of conjoint analysis and allows us to estimate individual-level differences using a multilevel model and determine both aggregate respondent preferences and the causal effect of each experimental feature on those preferences.

@@

We analyze the results using a multilevel Bayesian multinomial model (see the appendix for complete model details). Our experimental data has a natural hierarchical structure, with 3 questions nested inside 12 separate experimental tasks, nested inside each of the 1,016 respondents, which lends itself to multilevel modeling [@JensenMarbleScheve:2021]. Since it was impossible for every respondent to see every possible combination of all 12,000 experimental tasks, multilevel modeling allows us to pool together information about respondents with similar characteristics facing similar sets of choices. Moreover, using random respondent effects provides natural regularization and shrinkage for our estimates---experimental tasks that happened to appear more often due to chance will be accounted for and their frequency will not bias the overall causal effect. We define our model and priors in @eq-model.

@@

We do not include any respondent-level covariates beyond the treatment variables. Because this is an experimental design, any statistical confounding is accounted for during the process of randomization and covariates should have no systematic effect on treatment effects. We do not work with the raw results of the multinomial model directly. Given the conjoint design, we instead create a complete balanced grid of all 576 combinations of feature levels (2 transparency × 2 accountability × 3 government relationships × 4 organizations × 4 issues × 3 funding) and use the model to calculate predicted probabilities of choice selection for each combination of possible treatment values. We then collapse this set of predicted probabilities into estimated marginal means (EMMs) for specific features of interest while marginalizing or averaging over all other predicted variables [@LeeperHoboltTilley:2020; @BremerBurgisser:2023]. This marginalization process allows us to isolate the statistical effect of each feature in isolation. We include a complete table of model results in Table A5, along with a brief illustration of converting from regression coefficients to estimated marginal means.

@@

We provide visualizations of the full posterior distributions of each of the effects of interest, and we report two distributional summary statistics: (1) the posterior median, (2) credible intervals based on a 95% equal-tailed quantile interval. However, we are generally less concerned with the exact point estimates of our causal effects and instead focus on the direction and relative magnitude of their posterior distributions. For inference, we calculate the probability of direction ($p_d$), or the probability that the posterior AMCE is strictly positive or negative---it is the proportion of the posterior AMCE that is the sign of the median.

@@

```{r}
#| label: fig-transp-acc
#| fig-width: 6
#| fig-height: 2.5
#| out-width: "80%"
#| fig-cap: "Estimated marginal means for the four combinations of transparency and accountability."
#| include: true
mms_trans_acc_all <- mms_transp_acc %>% 
  filter(
    (feat_transp == "Transparency: Yes" & feat_acc == "Accountability: Yes") | 
      (feat_transp == "Transparency: No" & feat_acc == "Accountability: No")) %>% 
  bind_rows(
    mms_transp %>% 
      filter(feature_level == "Transparency: Yes") %>% 
      rename(feat_transp = feature_level) %>% 
      mutate(feat_acc = "Accountability: No"),
    mms_acc %>% 
      filter(feature_level == "Accountability: Yes") %>% 
      rename(feat_acc = feature_level) %>% 
      mutate(feat_transp = "Transparency: No")
  )

@@

# Discussion and conclusion

@@

These results have two important practical implications for NGO operations, fundraising, and survivability at a time when many INGOs are dealing with hostile host governments. First, while NGOs have little to no control over host government crackdowns over their organizations, they do have control over organizational practices such as transparency and accountability, and more importantly, public perception of these organizational characteristics. As a result, they may need to rely entirely on improving individual donor perceptions of organizational transparency and accountability and emphasizing the need for private donor funds at a time of shrinking civic space. This article thus provides insight into the importance of different framing or informational heuristics that can motivate such individual donor giving. In addition to voluntarily emphasizing or sharing this information with likely donors, nonprofits could also potentially brainstorm how to get more positive media exposure regarding these characteristics in restrictive civic environments. Moreover, we find that nonprofit transparency and accountability practices share the same causal effect, implying that donors might see these as interchangeable concepts.

Second, our results also show the importance of brand recognition and goodwill---respondents were more likely to prefer donating to an organization when it was identified as the Red Cross, a signal to the ubiquity of its brand across the globe and its emphasis on neutrality in hostile environments. Previous research shows that both brand awareness and brand trust are critical to non-involved consumers [@YenCaskeyWarden:2023]---in this case, donors who have previously not donated to an organization. Others show that donors are loyal to brands whose values are congruent to their own [@SargeantFordHudson:2008]. Our results thus have important implications on why nonprofits in hostile environments should focus on building brand awareness, as well as highlighting their values---to engage both likely and less likely donors.

Our results also have multiple important theoretical implications that should be explored in future research. First, future research should test how well these results map on to elite individual donors and foundations---not just the average private donor. As large foundations are more likely to publish data about their giving than the average private donor, it may be possible to determine how the changes in preferences elicited by these heuristics translate into actual donation behavior. Second, it is also important to test how these findings translate to non-American populations. Due to structural differences in the European nonprofit sector, where NGOs rely less on government funding and more on private funding [@Stroup:2012], donors are likely motivated by different concerns. Third, many states use negative rhetoric designed to sow mistrust between these groups and communities that support them. While donor responses have largely focused on navigating and adapting to legal restrictions, the use of such rhetoric raises broader questions about research and policy in the nonprofit sector: how is such rhetoric changing public attitudes and donations towards these groups? Are the effects of transparency and accountability found in this paper still likely to hold true if the home country also starts smearing these NGOs? Finally, though this paper examined legal crackdowns on INGOs, this issue is even more challenging for local NGOs in Global South countries, where individual donors might avoid giving to organizations focusing on contentious causes such as advocacy, media freedom, and anti-corruption initiatives due to unfavorable tax benefits or fear of retribution [@Brechenmacher:2017; @DupuyRonPrakash:2015; @Baoumi:2016]. Future research should examine how local NGOs can overcome individual foreign donors' concerns about host governments' criticisms of and crackdown on these groups.

# Data availability statement

All data and replication code is available at `DATAVERSE DOI` and at `OSF DOI`.

# Supplementary material

A companion statistical analysis notebook with a fuller description of the data, modeling details and diagnostics, and additional explanations of the Bayesian multinomial modeling approach, along with links to the data and replication code, is available at `ANONYMIZED URL/DOI`.

# Funding

@@

# Competing interests

The authors report there are no competing interests to declare.

::: {.content-visible when-profile="ms"}
```{=latex}
\newpage
\begingroup
\singlespacing
\parindent 0pt
\parskip 2ex
\def\enotesize{\normalsize}
\theendnotes
\endgroup
\newpage
```
:::

>>>>

---
title: "Navigating Hostility: The Effect of Nonprofit Transparency and Accountability on Donor Preferences in the Face of Shrinking Civic Space"
short-title: "Navigating Hostility"
date: "2025-04-14"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Utah State University
  url: https://occasionaldivergences.com/
  email: marc.dotson@usu.edu
  orcid: 0000-0001-7846-3660
- name: Andrew Heiss
  affiliations: 
    - id: gsu
      name: Georgia State University
      department: Andrew Young School of Policy Studies, Department of Public Management and Policy
      address: "55 Park Place NE, #464"
      city: Atlanta
      region: GA
      country: USA
      postal-code: 30303
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
  attributes:
    corresponding: true
correspondence-prefix: "Correspondence concerning this article should be addressed to"
abstract: |
  Governments across the world have increasingly used laws to restrict the work of nonprofits, which has led to a reduction in public or official foreign aid directed towards these groups. Many international nonprofits, in response, have turned to individual donors to offset the loss of traditional funding. What are individual donors' preferences regarding donating to legally besieged nonprofits abroad? We conducted a conjoint experiment on a nationally representative sample of likely donors in the US and found that learning about host government criticism and legal restrictions on nonprofits decreases individuals' preferences to donate to them. However, organizational features such as financial transparency and accountability can protect against this dampening effect. Our results have important implications both for understanding private international philanthropy and how nonprofits can better frame their fundraising appeals at a time when they are facing restrictive civic spaces and hostile governments abroad.
additional-info: |
  **Data Availability:** All data and replication code are publicly available at <https://github.com/andrewheiss/silent-skywalk>
keywords:
  - philanthropy
  - conjoint experiments
  - donor heuristics
  - repression
  - NGOs
  - civil society
  - nonprofits
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
csl: csl/apa.csl
word-count: "10,077"  # Subtract 197 for the final statements at the end
---

@@

tar_load(c(grid_treatment_only, level_lookup, feature_lookup, data_sans_conjoint))

details <- list(
  n = nrow(data_sans_conjoint),
  n_nice = label_comma()(nrow(data_sans_conjoint)),
  combos = 2 * 2 * 3 * 4 * 4 * 3
)

@@

We fielded a conjoint experiment to a nationally representative sample of `r details$n_nice` adults in the US. This experiment allows us to assess individual-level giving preferences among likely donors in the US towards NGOs abroad working in a variety of issue areas---human rights, environment, emergency response, and refugee relief.[^not-elite] More importantly, it allows us to directly compare the relative impact of host government restriction of NGOs and organizational financial transparency and accountability on individual donor preferences. The result of this experiment shows that both financial transparency and accountability increase individuals' willingness to donate to NGOs. Conversely, learning about host government criticism of and restrictions on NGOs decreases the likelihood of donation by itself; however, organizational features such as financial transparency and accountability can protect against this dampening effect.

@@

A conjoint research design also allows us to estimate treatment effects even if respondents do not see every combination of organization attributes, and correspondingly provides a substantial increase in statistical power. In standard factorial experimental designs, participants would ordinarily need to be shown all possible combinations of experimental attributes. In our experiment, we presented respondents with six possible treatments with randomized attributes for each treatment, yielding `r details$combos` possible unique combinations of features. Even with respondents answering twelve iterations of the experiment, not every combination was seen. However, as long as all the possible organizational attributes are well randomized and there are no systematic biases toward specific options (i.e., more respondents select the first option because it is the first) or toward earlier iterations of attribute choices (i.e., respondents are more careful and attentive for the first hypothetical organization than the last), we can pool all observations together for specific attributes of interest while marginalizing across all other attributes [@KertzerRenshonYarhi-Milo:2021]. This allows us to (1) estimate the effect of each experimental treatment even if some unique combinations of attributes were unseen, and (2) use a much smaller sample size than would be required in a more traditional factorial design.

While the ability to select key attributes and marginalize over others provides us with analytic flexibility, it also raises possible issues with multiple comparisons, *p*-hacking, and selective cherry-picking [@BansakHainmuellerHopkins:2021], especially given the fact that we have `r details$combos` possible combinations of independent variables to explore. As such, before launching the survey experiment, we preregistered a subset of confirmatory and exploratory hypotheses at the Open Science Framework.[^prereg] These hypotheses deal specifically with our key research questions about the effect of transparency, accountability, and crackdown on the propensity to donate, along with comparison treatment effects for organization brand name, issue area, and funding sources. We also examine the interaction between transparency, accountability, and host-country relationships.

[^prereg]: Our preregistration protocol is available at <https://doi.org/10.17605/OSF.IO/HSBYD> and in the appendix, and our data and reproducible code is available at <https://doi.org/10.17605/OSF.IO/R59XZ>. For clarity, we focus on our simplest, non-nested hypotheses, omitting many preregistered ones that overlap (e.g., the effect of branding and issue area and transparency and crackdown simultaneously). Following @WillrothAtherton:2024, we outline all the deviations from our preregistration protocol in the appendix.

@@

We fielded our survey experiment through Centiment, which recruits representative samples of paid (and highly engaged) survey participants online. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question about their philanthropic behavior early in the survey---if a participant responded that they give once every few years or never, they were disqualified from the study. We also included an attention check question early in the survey and removed respondents who failed the question. Importantly, these screening questions were presented *prior to* the experimental manipulation to avoid post-treatment bias [@MontgomeryNyhanTorres:2018]. Following screening, we received `r details$n_nice` viable responses. Tables A1, A2, and A3 provide a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across all pre-treatment characteristics, including gender, age, education, income, and attitudes toward charity. Additionally, Table A4 compares sample characteristics with nationally representative estimates from the 2019 Current Population Survey (CPS). Our sample matches national proportions of age, marital status, and education, and contains slightly more men and slightly higher income levels than the general population. Due to the initial screening question that limited participants to those with an interest in charitable activity, respondents are far more likely to have donated, volunteered, or voted in the past year than population CPS estimates. Our reported causal effects are therefore not representative of the entire population and instead are generalizable to people interested in charitable giving.

@@

At the time of designing the study, Charity Navigator---a large database of US nonprofit financial information---categorized international nonprofit activities into four general causes: (1) humanitarian relief supplies, (2) international peace, security, and affairs, (3) development and relief services, and (4) environment. Accordingly, we used four nonprofits that are stereotypical for each cause: the International Committee of the Red Cross (humanitarian relief), Amnesty International (human rights), Oxfam (development), and Greenpeace (environment).[^real-names] We also varied several other features, including four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), two organizational practices (financial transparency and third-party accountability audits), three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and three relationships with host governments (friendly, criticized by the government, and under government crackdown).

[^cn]: See <http://web.archive.org/web/20170315000746/https://www.charitynavigator.org/index.cfm?bay=search.categories&categoryid=7>. At the time of writing, Charity Navigator had changed its taxonomy to include 18 broader categories.

@@

After presenting respondents with a set of three randomly generated organizations, we asked them which of the three they would be willing to donate to, along with an option for no selection (see @tbl-example for an example). This choice indicated donation intention and reflected the option the respondent preferred most within that set. We then repeated the process eleven more times for a total of twelve randomized iterations of hypothetical combinations of attributes, resulting in `r label_comma()(details$n * 12)` completed experimental tasks (12 iterations × `r details$n_nice` respondents). This iterative process across twelve experimental tasks is typical of conjoint analysis and allows us to estimate individual-level differences using a multilevel model and determine both aggregate respondent preferences and the causal effect of each experimental feature on those preferences.

@@

We analyze the results using a multilevel Bayesian multinomial model (see the appendix for complete model details). Our experimental data has a natural hierarchical structure, with 3 questions nested inside 12 separate experimental tasks, nested inside each of the `r details$n_nice` respondents, which lends itself to multilevel modeling [@JensenMarbleScheve:2021]. Since it was impossible for every respondent to see every possible combination of all 12,000+ experimental tasks, multilevel modeling allows us to pool together information about respondents with similar characteristics facing similar sets of choices. Moreover, using random respondent effects provides natural regularization and shrinkage for our estimates---experimental tasks that happened to appear more often due to chance will be accounted for and their frequency will not bias the overall causal effect. We define our model and priors in @eq-model.

@@

We do not include any respondent-level covariates beyond the treatment variables. Because this is an experimental design, any statistical confounding is accounted for during the process of randomization and covariates should have no systematic effect on treatment effects. We do not work with the raw results of the multinomial model directly. Given the conjoint design, we instead create a complete balanced grid of all `r details$combos` combinations of feature levels (2 transparency × 2 accountability × 3 government relationships × 4 organizations × 4 issues × 3 funding) and use the model to calculate predicted probabilities of choice selection for each combination of possible treatment values. We then collapse this set of predicted probabilities into estimated marginal means (EMMs) for specific features of interest while marginalizing or averaging over all other predicted variables [@LeeperHoboltTilley:2020; @Arel-BundockGreiferHeiss:2024]. This marginalization process allows us to isolate the statistical effect of each feature in isolation. We include a complete table of model results in Table A5, along with a brief illustration of converting from regression coefficients to estimated marginal means.

@@

We provide visualizations of the full posterior distributions of each of the effects of interest, and we report two distributional summary statistics: (1) the posterior median, (2) credible intervals based on a 95% equal-tailed quantile interval. However, we are generally less concerned with the exact point estimates of our causal effects and instead focus on the direction and relative magnitude of their posterior distributions. For inference, we calculate the probability of direction ($p_d$), which indicates the proportion of the posterior AMCE that shares the sign of the median and the probability that AMCE is strictly positive or negative.

@@

```{r}
#| label: fig-transp-acc
#| fig-width: 6
#| fig-height: 2.5
#| out-width: "80%"
#| fig-cap: "Estimated marginal means for the four combinations of transparency and accountability"
#| include: true
mms_trans_acc_all <- mms_transp_acc %>% 
  filter(
    (feat_transp == "Transparency: Yes" & feat_acc == "Accountability: Yes") | 
      (feat_transp == "Transparency: No" & feat_acc == "Accountability: No")) %>% 
  bind_rows(
    mms_transp %>% 
      filter(feature_level == "Transparency: Yes") %>% 
      rename(feat_transp = feature_level) %>% 
      mutate(feat_acc = "Accountability: No"),
    mms_acc %>% 
      filter(feature_level == "Accountability: Yes") %>% 
      rename(feat_acc = feature_level) %>% 
      mutate(feat_transp = "Transparency: No")
  )

@@

# Discussion

These results have two important practical implications for NGO operations, fundraising, and survivability at a time when many INGOs are dealing with hostile host governments. First, while NGOs have little to no control over the legal environments of their host governments, they do have control over organizational practices such as transparency and accountability, and more importantly, public perception of these organizational characteristics. As a result, they may need to rely entirely on improving individual donor perceptions of organizational transparency and accountability and emphasizing the need for private donor funds at a time of shrinking civic space. These results provide insight into the importance of different framing or informational heuristics that can motivate such individual donor giving. In addition to voluntarily emphasizing or sharing this information with likely donors, nonprofits could also potentially brainstorm how to get more positive media exposure regarding these characteristics in restrictive civic environments. Moreover, we find that nonprofit transparency and accountability practices share the same causal effect, implying that donors might see these as interchangeable concepts.

Second, our results also highlight the importance of brand recognition and goodwill---respondents were more likely to prefer donating to an organization when it was identified as the Red Cross, a signal to the ubiquity of its brand across the globe and its emphasis on neutrality in hostile environments. Previous research shows that both brand awareness and brand trust are critical to non-involved consumers [@YenCaskeyWarden:2023]---in this case, donors who have previously not donated to an organization. Others show that donors are loyal to brands whose values are congruent to their own [@SargeantFordHudson:2008]. Our results thus have important implications on why nonprofits in hostile environments should focus on building brand awareness, as well as highlighting their values---to engage both likely and less likely donors.

# Conclusion

@@

Our results also have multiple important theoretical implications that should be explored in future research. First, future research should test how well these results map on to elite individual donors and foundations---not just the average private donor. As large foundations are more likely to publish data about their giving than the average private donor, it may be possible to determine how the changes in preferences elicited by these heuristics translate into actual donation behavior. Second, it is also important to test how these findings translate to non-American populations. Due to structural differences in the European nonprofit sector, where NGOs rely less on government funding and more on private funding [@Stroup:2012], donors are likely motivated by different concerns. Third, many states use negative rhetoric designed to sow mistrust between these groups and communities that support them. While donor responses have largely focused on navigating and adapting to legal restrictions, the use of such rhetoric raises broader questions about research and policy in the nonprofit sector: how is such rhetoric changing public attitudes and donations towards these groups? Are the effects of transparency and accountability found in this paper still likely to hold true if the home country also starts smearing these NGOs? Finally, though this paper examined legal crackdowns on INGOs, this issue is even more challenging for local NGOs in Global South countries, where individual donors might avoid giving to organizations focusing on contentious causes such as advocacy, media freedom, and anti-corruption initiatives due to unfavorable tax benefits or fear of retribution [@Brechenmacher:2017; @DupuyRonPrakash:2015]. Future research should examine how local NGOs can overcome individual foreign donors' concerns about host governments' criticisms of and crackdown on these groups.

# Statements and declarations

## Data availability

All data and replication code is available at <https://doi.org/10.5281/zenodo.14721277> and at <https://doi.org/10.17605/OSF.IO/R59XZ>.

We prereigstered our hypotheses and research design at the Open Science Framework (OSF). The full version of the OSF preregistration protocol is available at <https://doi.org/10.17605/OSF.IO/HSBYD>; a condensed version is included in the online appendix. We also include a list of deviations from the preregistration in the online appendix.

## Ethical considerations

This experiment was approved and deemed as exempt by the human subjects research committees at both Christopher Newport University (1436622-1), Brigham Young University (E19135), and Georgia State University (H19644).

## Consent to participate

We obtained written informed consent from each survey participant. The full text of the consent statement is included in the online appendix.

## Consent for publication

Not applicable.

## Declaration of conflicting interest

The authors declare no potential conflicts of interest with respect to the research, authorship, and/or publication of this article.

## Funding

@@

## Supplementary material

A companion statistical analysis notebook with a more complete description of the data, modeling details and diagnostics, and additional explanations of the Bayesian multinomial modeling approach, along with links to the data and replication code, is available at <https://stats.andrewheiss.com/silent-skywalk/>.
